---
title: "Getting started with happifn"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting started with happifn}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(happifn)
```

# Get raw datasets

`happifn` can be used to retrieve raw data from permanent sample plots. Seven datasets are available. Their names can be retrieved using `get_datasets_names()` : 

```{r get_dataset_names}
get_dataset_names()
```

Then simply use the `get_ifn()` function to retrieve them. If no directory is specified, the data is seamlessly loaded into R from a url. If you wanna use a directory, ".zip" file from `get_url()` must be downloaded upstream and then unzipped in the same directory. 

**Note:** The `placette` dataset will be automatically converted to a spatial `sf` object. 

```{r get_ifn, eval=FALSE}
arbre <- get_ifn("arbre", dir = NULL)
```

There a wrapper function to get all data at once : `get_ifn_all()`. That will download and load all datasets in one go. 

```{r get_ifn_all, eval=FALSE}
get_ifn_all()
```

For a better understanding of how datasets are constructed, you can retrieve the metadata using `get_ifn_metadata()`. Three metadatasets are availables : "code", "units", "units_value_set". If `unlist = TRUE`, instead of having a list with three tables, each table will be added to the environment as a separate variable. 


```{r get_ifn_metadata, eval=FALSE}
metadata <- get_ifn_metadata()

code <- metadata[[1]]
```

# Get spatial datasets

The IFN also provides geographic layers grouping together different forest regions. As these layers are heavy, they are stored in another package. To load them, use `check_happifndata()`. For the moment, four data sets are available :  

- "ser" : sylvoecoregion; 
- "ser_ar" : sylvoecoregion d'alluvions recents; 
- "rfn" : regions forestieres nationales 
- "zp" : zone phytogeographiques 

```{r happifndata, eval = FALSE}
check_happifndata()
library(happifndata)
 
data("ser")
data("ser_ar")
data("rfn")
data("zp")

```

